---
title: "Datasets"
author: "Luke Zappia"
date: '`r Sys.Date()`'
output: html_document
---

Code version: `r system("git log -1 --format=oneline | cut -d' ' -f1", intern = TRUE)`

```{r knitr, include = FALSE}
DOCNAME = "datasets"
knitr::opts_chunk$set(autodep        = TRUE,
                      cache          = TRUE,
                      cache.path     = paste0("cache/", DOCNAME, "/"),
                      cache.comments = TRUE,
                      echo           = TRUE,
                      error          = FALSE,
                      fig.align      = "center",
                      fig.path       = paste0("figures/", DOCNAME, "/"),
                      fig.width      = 10,
                      fig.height     = 8,
                      dev            = c('svg', 'png'),
                      message        = FALSE,
                      warning        = FALSE)
```

```{r libraries, cache = FALSE}
# RNA-seq
library("scater")
library("splatter")

# Parallel
library("BiocParallel")

# Plotting
library("cowplot")

# Tables
library("knitr")

# Tidyverse
library("tidyverse")
```

```{r source, cache = FALSE}
source("../R/load_datasets.R")
source("../R/simulate_datasets.R")
source("../R/test_genes.R")
source("../R/utils.R")
```

```{r ggtheme, cache = FALSE}
```

Introduction
============

In this document we are going to load multiple real datasets, estimate
parameters for the various simulations, simulate new datasets and compare them
to the real data.

Data
====

We have a variety of real datasets we could use. Let's load the metadata file
and take a look.

```{r datasets}
root <- "../data"
datasets <- read_tsv(file.path(root, "datasets.txt"),
                     col_types = cols(.default = col_character(),
                                      NumCells = col_integer()
                                      )
                     )
datasets
```

These five datasets cover a variety of different species, cell types, platforms,
processing etc.

For the paper we need a table, which I am going to save in LaTeX format.

```{r datasets-table}
datasets %>%
    select(Dataset, Species, CellType, Platform, Protocol, UMI, NumCells) %>%
    arrange(Dataset) %>%
    kable(format = "latex", caption = "Details of real datasets") %>%
    write_lines("../output/datasets_table.tex")
```

Simulation
==========

The next step is to load in the datasets, estimate parameters from them,
simulate using the various models in Splatter and compare the results. Here we
do that for each dataset. We also produce an overall comparison panel.

```{r simulate, cache.lazy = FALSE}
bp <- BiocParallel::MulticoreParam(10)

res <- list()

for (idx in seq_len(nrow(datasets))) {
    dataset <- datasets[idx, ]
    message("#### STARTING ", dataset["Dataset"], " ####")
    message("***LOADING COUNTS***")
    counts <- loadDataset(dataset, root)
    counts <- counts[rowSums(counts) > 0, ]
    na.rows <- which(rowSums(is.na(counts)) > 0)
    if (length(na.rows) > 0) {
        counts <- counts[-na.rows, ]
    }
    counts <- counts[, sample(1:ncol(counts), 200)]
    name <- unlist(dataset["Dataset"])
    comp <- simComp(dataset, root, bp = bp)
    res[[name]] <- comp
    message("Writing comparison...")
    write_rds(comp, path = paste0("../output/res_", name, ".Rds"),
              compress = "gz")
    message("All done!")
}
```

```{r make-overall, cache.lazy=FALSE}
for (idx in seq_along(res)) {
    name <- names(res)[idx]
    message(paste("Making", name, "overall panel..."))
    comp <- res[[name]]
    plt <- makeOverallPanel(comp$Comp, comp$Diff, title = name)
    plt.name <- paste0("additional", idx + 3, "_", name, "_overall")
    save_plot(paste0("../additional/", plt.name, ".png"), plt,
              ncol = 4, nrow = 7, base_height = 3)
    #save_plot(paste0("../additional/", plt.name, ".pdf"), plt,
    #          ncol = 4, nrow = 7, base_height = 3)
}
```

Tung comparison
===============

The overall comparison panels are good for getting an idea of how simulations
perform but the format is not really appropriate for a paper. We are going to
extract the plots for the Tung dataset, pretty them up a bit and put them into
paper appropriate panels.

```{r tung-panel}
plots <- list(MeansComp   = res$Tung$Comp$Plots$Means,
              MeansDiff   = res$Tung$Diff$Plots$Means,
              VarsComp    = res$Tung$Comp$Plots$Variances,
              VarsDiff    = res$Tung$Diff$Plots$Variances,
              MeanVarComp = res$Tung$Comp$Plots$MeanVar,
              MeanVarDiff = res$Tung$Diff$Plots$MeanVar,
              LibSizeComp = res$Tung$Comp$Plots$LibrarySizes,
              LibSizeDiff = res$Tung$Diff$Plots$LibrarySizes)

# Select the colour we are going to use
cols <- RColorBrewer::brewer.pal(8, "Dark2")

# Labels for datasets
labels = c("Real" = "Real", "Splat" = "Splat", "SplatDrop" = "Splat (Drop)",
           "Simple" = "Simple", "Lun" = "Lun", "Lun2" = "Lun 2",
           "Lun2ZINB" = "Lun 2 (ZINB)")

# Make adjustments to the plots
for (idx in seq_along(plots)) {
    name <- names(plots)[idx]
    plot <- plots[[idx]]
    
    # Set a new theme
    plot <- plot +
        theme_cowplot(font_size = 12) +
        theme(legend.position = "none")
    
    # Set the colours, diff plots have one less dataset
    if (grepl("Comp", name)) {
        plot <- plot + scale_color_manual(values = cols, labels = labels) + 
            scale_fill_manual(values = cols, labels = labels)
    } else {
        plot <- plot + scale_color_manual(values = cols[-1], labels = labels[-1]) + 
            scale_fill_manual(values = cols[-1], labels = labels[-1])
    }
    
    # Boxplots are replotted with different properties, axis text adjusted and
    # x label removed
    if (!grepl("MeanVar", name)) {
        plot <- plot + geom_boxplot(aes(fill = Dataset),
                                    size = 1.5, alpha = 0.2) +
            scale_x_discrete(labels = c("Splat" = "Splat",
                                        "SplatDrop" = "Splat (Drop)",
                                        "Simple" = "Simple", "Lun" = "Lun",
                                        "Lun2" = "Lun 2",
                                        "Lun2ZINB" = "Lun 2 (ZINB)")) +
            theme(axis.title.x = element_blank(),
                  axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))
    }

    # Set y-axis on some plots
    if (name == "VarsComp") {
        plot <- plot + coord_cartesian(ylim = c(0, 20))
    }
    if (name == "VarsDiff") {
        plot <- plot + coord_cartesian(ylim = c(-10, 20))
    }
    if (name == "MeanVarComp") {
        plot <- plot + coord_cartesian(ylim = c(0, 30))
    }
    if (name == "MeanVarDiff") {
        plot <- plot + coord_cartesian(ylim = c(-10, 30))
    }
    
    plots[[idx]] <- plot
}

# Get a single legend to use
leg <- get_legend(plots[["MeanVarComp"]] + theme(legend.position = "bottom"))

# Assemble the panel
panel <- ggdraw() +
    draw_label("A", 0.01, 0.986,
               fontface = "bold", hjust = 0, vjust = 0) +
    draw_plot(plots$MeansComp,  0.00, 0.77, 0.49, 0.23) +
    draw_label("B", 0.51, 0.986,
               fontface = "bold", hjust = 0, vjust = 0) +
    draw_plot(plots$MeansDiff,  0.51, 0.77, 0.49, 0.23) +
    draw_label("C", 0.01, 0.746,
               fontface = "bold", hjust = 0, vjust = 0) +
    draw_plot(plots$VarsComp,  0.00, 0.53, 0.49, 0.23) +
    draw_label("D", 0.51, 0.746,
               fontface = "bold", hjust = 0, vjust = 0) +
    draw_plot(plots$VarsDiff,  0.51, 0.53, 0.49, 0.23) +
    draw_label("E", 0.01, 0.506,
               fontface = "bold", hjust = 0, vjust = 0) +
    draw_plot(plots$MeanVarComp,  0.00, 0.29, 0.49, 0.23) +
    draw_label("F", 0.51, 0.506,
               fontface = "bold", hjust = 0, vjust = 0) +
    draw_plot(plots$MeanVarDiff,  0.51, 0.29, 0.49, 0.23) +
    draw_label("G", 0.01, 0.266,
               fontface = "bold", hjust = 0, vjust = 0) +
    draw_plot(plots$LibSizeComp,  0.00, 0.05, 0.49, 0.23) +
    draw_label("H", 0.51, 0.266,
               fontface = "bold", hjust = 0, vjust = 0) +
    draw_plot(plots$LibSizeDiff, 0.51, 0.05, 0.49, 0.23) +
    draw_plot(leg, 0.00, 0.00, 1.00, 0.04)

save_plot("../figures/figure2_tung_comparison.png", panel, ncol = 2, nrow = 4)
save_plot("../figures/figure2_tung_comparison.pdf", panel, ncol = 2, nrow = 4)
panel
```

```{r tung-panel-zeros}
plots <- list(ZerosCellComp = res$Tung$Comp$Plots$ZerosCell,
              ZerosCellDiff = res$Tung$Diff$Plots$ZerosCell,
              ZerosGeneComp = res$Tung$Comp$Plots$ZerosGene,
              ZerosGeneDiff = res$Tung$Diff$Plots$ZerosGene,
              MeanZerosComp = res$Tung$Comp$Plots$MeanZeros,
              MeanZerosDiff = res$Tung$Diff$Plots$MeanZeros)

# Make adjustments to the plots
for (idx in seq_along(plots)) {
    name <- names(plots)[idx]
    plot <- plots[[idx]]
    
    # Set a new theme
    plot <- plot +
        theme_cowplot(font_size = 12) +
        theme(legend.position = "none")
    
    # Set the colours, diff plots have one less dataset
    if (grepl("Comp", name)) {
        plot <- plot + scale_color_manual(values = cols, labels = labels) + 
            scale_fill_manual(values = cols, labels = labels)
    } else {
        plot <- plot + scale_color_manual(values = cols[-1], labels = labels[-1]) + 
            scale_fill_manual(values = cols[-1], labels = labels[-1])
    }
    
    # Boxplots are replotted with different properties, axis text adjusted and
    # x label removed
    if (!grepl("MeanZeros", name)) {
        plot <- plot + geom_boxplot(aes(fill = Dataset),
                                    size = 1.5, alpha = 0.2) +
            scale_x_discrete(labels = c("Splat" = "Splat",
                                        "SplatDrop" = "Splat (Drop)",
                                        "Simple" = "Simple", "Lun" = "Lun",
                                        "Lun2" = "Lun 2",
                                        "Lun2ZINB" = "Lun 2 (ZINB)")) +
            theme(axis.title.x = element_blank(),
                  axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))
    }
    
    plots[[idx]] <- plot
}

# Get a single legend to use
leg <- get_legend(plots[["MeanZerosComp"]] + theme(legend.position = "bottom"))

# Assemble the panel
panel <- ggdraw() +
    draw_label("A", 0.01, 0.982,
               fontface = "bold", hjust = 0, vjust = 0) +
    draw_plot(plots$ZerosCellComp,  0.00, 0.69, 0.49, 0.31) +
    draw_label("B", 0.51, 0.982,
               fontface = "bold", hjust = 0, vjust = 0) +
    draw_plot(plots$ZerosCellDiff,  0.51, 0.69, 0.49, 0.31) +
    draw_label("C", 0.01, 0.662,
               fontface = "bold", hjust = 0, vjust = 0) +
    draw_plot(plots$ZerosGeneComp,  0.00, 0.37, 0.49, 0.31) +
    draw_label("D", 0.51, 0.662,
               fontface = "bold", hjust = 0, vjust = 0) +
    draw_plot(plots$ZerosGeneDiff,  0.51, 0.37, 0.49, 0.31) +
    draw_label("E", 0.01, 0.342,
               fontface = "bold", hjust = 0, vjust = 0) +
    draw_plot(plots$MeanZerosComp,  0.00, 0.05, 0.49, 0.31) +
    draw_label("F", 0.51, 0.342,
               fontface = "bold", hjust = 0, vjust = 0) +
    draw_plot(plots$MeanZerosDiff,  0.51, 0.05, 0.49, 0.31) +
    draw_plot(leg, 0.00, 0.00, 1.00, 0.04)

save_plot("../figures/figure3_tung_zeros.png", panel, ncol = 2, nrow = 3)
save_plot("../figures/figure3_tung_zeros.pdf", panel, ncol = 2, nrow = 3)
panel
```

Overall summary
===============

We have produced plots to show how simulations compare to individual datasets
but what about an overall summary across all of them? To make this easier we
are going to calculate a single statistic for each of the different properties
we are interested in using the `summariseDiff` function. For each property the
values for both the real data and the simulations are sorted, the difference
between them calculated and the Median Absolute Deviation is taken as the
statistic.

```{r summarise}
summary <- list()

for (name in names(res)) {
    diff <- res[[name]]$Diff
    summ <- summariseDiff(diff)$Long
    colnames(summ)[1] <- "Simulation"
    summ <- data.frame(Dataset = name, summ)
    summary[[name]] <- summ
}

summary <- bind_rows(summary)
```

In the paper we are going to present these results as a heatmap of the ranks.

```{r summary-heatmap}
summary %>%
    mutate(Statistic = factor(Statistic,
                              levels = rev(c("Mean", "Variance", "MeanVar",
                                         "LibSize", "ZerosCell", "ZerosGene",
                                         "MeanZeros")))) %>%
    mutate(Rank = factor(Rank)) %>%
    ggplot(aes(x = Simulation, y = Statistic, fill = Rank, label = MAD)) +
    geom_tile() +
    viridis::scale_fill_viridis(direction = -1, discrete = TRUE) +
    scale_x_discrete(labels = c("Splat" = "Splat",
                                "SplatDrop" = "Splat (Drop)",
                                "Simple" = "Simple", "Lun" = "Lun",
                                "Lun2" = "Lun 2",
                                "Lun2ZINB" = "Lun 2 (ZINB)")) +
    scale_y_discrete(labels = c("Mean" = "Mean", "Variance" = "Variance",
                                "MeanVar" = "Mean-Variance",
                                "LibSize" = "Library Size",
                                "ZerosCell" = "% Zeros (Cell)",
                                "ZerosGene" = "% Zeros (Gene)",
                                "MeanZeros" = "Mean-Zeros")) +
    ggtitle("Rank of MAD from real data") +
    guides(fill = guide_legend(ncol = 1)) +
    facet_grid(Dataset ~ .) +
    theme_minimal() +
    theme(plot.title = element_text(size = 24, face = "bold", hjust = 0.4),
          axis.title = element_blank(),
          axis.text = element_text(size = 12),
          axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5),
          panel.grid = element_blank(),
          strip.text = element_text(size = 16, face = "bold"),
          legend.title = element_text(size = 14),
          legend.text = element_text(size = 12),
          legend.key.size = unit(25, "points"),
          legend.position = "right")

ggsave("../figures/figure4_mads_heatmap.pdf",
       width = 20, height = 25, units = "cm")
ggsave("../figures/figure4_mads_heatmap.png",
       width = 20, height = 25, units = "cm")
```

We also produce a supplementary table of the MADs as a CSV.

```{r write-summary}
write_csv(summary, "../additional/additional12_mads.csv")
```

Gene goodness-of-fit
====================

```{r goodness-of-fit}
res$Comp$FeatureData %>%
    rowwise() %>%
    mutate(IsNA = all(is.na(NBChi), is.na(LNChi), is.na(NormChi))) %>%
    #filter(!IsNA) %>%
    mutate(IsNB = NBPVal > 0.01, IsLN = LNPVal > 0.01,
           IsNorm = NormPVal > 0.01) %>%
    ungroup() %>%
    group_by(Dataset) %>%
    summarise(NotFit = sum(IsNA), NB = sum(IsNB, na.rm = TRUE),
              LNorm = sum(IsLN, na.rm = TRUE),
              Norm = sum(IsNorm, na.rm = TRUE)) %>%
    gather(key = Dist, value = Count, -Dataset) %>%
    ggplot(aes(x = Dataset, y = Count, fill = Dist)) +
    geom_bar(stat = "identity", width = 0.4, 
             position = position_dodge(width = 0.5))
```

Session info
============

```{r session-info, cache = FALSE}
devtools::session_info()
```
