---
title: "Performance"
author: "Luke Zappia"
date: '`r Sys.Date()`'
output: html_document
---

Code version: `r system("git log -1 --format=oneline | cut -d' ' -f1", intern = TRUE)`

```{r knitr, include = FALSE}
DOCNAME = "performance"
knitr::opts_chunk$set(autodep        = TRUE,
                      cache          = TRUE,
                      cache.path     = paste0("cache/", DOCNAME, "/"),
                      cache.comments = TRUE,
                      echo           = TRUE,
                      error          = FALSE,
                      fig.align      = "center",
                      fig.path       = paste0("figures/", DOCNAME, "/"),
                      fig.width      = 10,
                      fig.height     = 8,
                      dev            = c('svg', 'png'),
                      message        = FALSE,
                      warning        = FALSE)
```

```{r libraries, cache = FALSE}
# RNA-seq
library("scater")
library("splatter")

# Parallel
library("BiocParallel")

# Plotting
library("cowplot")

# Tables
library("knitr")

# Tidyverse
library("tidyverse")
```

```{r source, cache = FALSE}
source("../R/load_datasets.R")
source("../R/simulate_datasets.R")
```

```{r ggtheme, cache = FALSE}
```

Introduction
============

In this document we are going to look at how the estimation and simulation
procedures perform as the number of genes and cells increases.

Data
====

We have a variety of real datasets we could use. Let's load the metadata file
and take a look.

```{r datasets}
root <- "../data"
datasets <- read_tsv(file.path(root, "datasets.txt"),
                     col_types = cols(.default = col_character(),
                                      NumCells = col_integer()
                                      )
                     )
datasets
```

For this experiment we are going to use just the Tung dataset.

```{r dataset}
dataset <- datasets[datasets$Dataset == "Tung", ]
counts <- loadDataset(dataset, root)
```

Genes
=====

Let's have a look at how the performance of the different models is affected
by an increasing number of genes with 100 cells.

```{r counts-100}
# Select 100 random cells
counts.100 <- counts[, sample(1:ncol(counts), 100)]
counts.100 <- counts.100[rowSums(counts.100) > 0, ]
```

```{r genes}
bp <- BiocParallel::MulticoreParam(10)

res <- bplapply(c(500, 1000, 1500, 2000), function(nGenes) {
    message("#### ", Sys.time(), " STARTING ", nGenes, " genes ####")
    set.seed(1)
    counts.100.nGenes <- counts.100[sample(1:nrow(counts.100), nGenes), ]
    times <- simComp(counts.100.nGenes, compare = FALSE,
                     test.gof = FALSE, verbose = TRUE)$Timings
    times <- times %>%
        data.frame() %>%
        rownames_to_column("Model") %>%
        mutate(nGenes = nGenes)
    message("#### ", Sys.time(), " All done! ####")
    return(times)
}, BPPARAM = bp)

timings <- bind_rows(res)
```

```{r plot-genes}
# Select the colours we are going to use
cols <- RColorBrewer::brewer.pal(9, "Set1")

# Labels for datasets
labels = c("Real" = "Real", "Splat" = "Splat", "SplatDrop" = "Splat (Drop)",
           "Simple" = "Simple", "Lun" = "Lun", "Lun2" = "Lun 2",
           "Lun2ZINB" = "Lun 2 (ZINB)", "scDD" = "scDD", "BASiCS" = "BASiCS")

timings %>%
    mutate(Model = factor(Model, levels = c("Splat", "SplatDrop", "Simple",
                                            "Lun", "Lun2", "Lun2ZINB", "scDD",
                                            "BASiCS"))) %>%
    gather(key = "Type", value = "Time", -Model, -nGenes) %>%
    filter(!(Model == "SplatDrop" & Type == "Estimation")) %>%
    filter(!(Model == "Lun2ZINB" & Type == "Estimation")) %>%
    ggplot(aes(x = nGenes, y = Time, colour = Model)) +
    geom_line() +
    geom_point() +
    scale_colour_manual(values = cols[-1], labels = labels[-1]) +
    scale_y_log10(labels = scales::comma) +
    facet_wrap(~ Type, nrow = 2, scales = "free") +
    ggtitle("Processing times (genes)") +
    xlab("Number of genes") +
    ylab("Time (seconds)") +
    theme(plot.title = element_text(size = 24, face = "bold", hjust = 0.4),
          strip.background = element_rect(fill = "white"),
          strip.text = element_text(size = 20),
          legend.position = "bottom")
```

```{r save-genes}
write_tsv(timings, "../output/performance_genes.tsv")
```

Cells
=====

We can do the same thing with increasing numbers of cells. Here we will use
1000 genes.

```{r cells}
res <- bplapply(c(50, 100, 150, 200), function(nCells) {
    message("#### ", Sys.time(), " STARTING ", nCells, " genes ####")
    set.seed(1)
    counts.nCells <- counts[, sample(1:ncol(counts), nCells)]
    counts.nCells <- counts.nCells[rowSums(counts.nCells) > 0, ]
    counts.nCells.1k <- counts.nCells[sample(1:nrow(counts.nCells), 1000), ]
    times <- simComp(counts.nCells.1k, compare = FALSE,
                     test.gof = FALSE, verbose = TRUE)$Timings
    times <- times %>%
        data.frame() %>%
        rownames_to_column("Model") %>%
        mutate(nCells = nCells)
    message("#### ", Sys.time(), " All done! ####")
    return(times)
}, BPPARAM = bp)

timings <- bind_rows(res)
```

```{r plot-cells}
timings %>%
    mutate(Model = factor(Model, levels = c("Splat", "SplatDrop", "Simple",
                                            "Lun", "Lun2", "Lun2ZINB", "scDD",
                                            "BASiCS"))) %>%
    gather(key = "Type", value = "Time", -Model, -nCells) %>%
    filter(!(Model == "SplatDrop" & Type == "Estimation")) %>%
    filter(!(Model == "Lun2ZINB" & Type == "Estimation")) %>%
    ggplot(aes(x = nCells, y = Time, colour = Model)) +
    geom_line() +
    geom_point() +
    scale_colour_manual(values = cols[-1], labels = labels[-1]) +
    scale_y_log10(labels = scales::comma) +
    facet_wrap(~ Type, nrow = 2, scales = "free") +
    ggtitle("Processing times (cells)") +
    xlab("Number of cells") +
    ylab("Time (seconds)") +
    theme(plot.title = element_text(size = 24, face = "bold", hjust = 0.4),
          strip.background = element_rect(fill = "white"),
          strip.text = element_text(size = 20),
          legend.position = "bottom")
```

```{r save-cells}
write_tsv(timings, "../output/performance_cells.tsv")
```

Session info
============

```{r session-info, cache = FALSE}
devtools::session_info()
```
