---
title: "Simulations"
author: "Luke Zappia"
date: '`r Sys.Date()`'
output: html_document
---

Code version: `r system("git log -1 --format=oneline | cut -d' ' -f1", intern = TRUE)`

```{r knitr, include = FALSE}
DOCNAME = "simulations"
knitr::opts_chunk$set(autodep        = TRUE,
                      cache          = TRUE,
                      cache.path     = paste0("cache/", DOCNAME, "/"),
                      cache.comments = TRUE,
                      echo           = FALSE,
                      error          = FALSE,
                      fig.align      = "center",
                      fig.path       = paste0("figures/", DOCNAME, "/"),
                      fig.width      = 10,
                      fig.height     = 8,
                      dev            = c('svg', 'png'),
                      message        = FALSE,
                      warning        = FALSE)
```

```{r libraries, cache = FALSE}
# RNA-seq
library("scater")
library("splatter")

# Plotting
library("cowplot")

# Tidyverse
library("tidyverse")
```

```{r source, cache = FALSE}
source("../R/utils.R")
```

```{r ggtheme, cache = FALSE}
my.pal <- mcriPalette("themes")
my.pal[4] <- "#F7DC5F"
```

Introduction
============

In this document we are going to demonstrate some of the types of simulations
possible using the `Splat` simulation in `Splatter`.

Different situations
==================

Single
------

Simulation of a single homogenous population of cells.

```{r single}
sim.single <- splatSimulateSingle(groupCells = 200, dropout.present = FALSE)
sim.single <- plotPCA(sim.single, return_SCESet = TRUE, draw_plot = FALSE)
```

```{r single-PCA}
data.frame(PC1 = reducedDimension(sim.single)[, 1],
           PC2 = reducedDimension(sim.single)[, 2]) %>%
    ggplot(aes(x = PC1, y = PC2)) +
    geom_point(size = 4, colour = "#092F5E")
```

Groups
------

Simulation of a sample with multiple types of cells.

```{r groups}
sim.groups <- splatSimulateGroups(groupCells = c(200, 100, 80, 20, 300, 120),
                                  de.prob = c(0.05, 0.05, 0.08, 0.05, 0.06,
                                              0.05), de.facLoc = 0.1,
                                  de.facScale = 0.4,
                                  dropout.present = TRUE,
                                  seed = 1)
sim.groups <- plotPCA(sim.groups, return_SCESet = TRUE, draw_plot = FALSE)
```

```{r groups-PCA}
pca.groups <- data.frame(PC1 = reducedDimension(sim.groups)[, 1],
           PC2 = reducedDimension(sim.groups)[, 2],
           group = pData(sim.groups)$Group) %>%
    ggplot(aes(x = PC1, y = PC2, colour = group)) +
    geom_point(size = 1.5) +
    scale_color_manual(values = my.pal, name = "Group") +
    guides(col = guide_legend(ncol = 1)) +
    ggtitle("Simulation with groups") +
    theme_cowplot() +
    theme(plot.title = element_text(size = 20, hjust = 0.4),
          axis.text = element_text(size = 12),
          legend.title = element_text(size = 14),
          legend.text = element_text(size = 12),
          legend.key.size = unit(25, "points"),
    )

pca.groups
```

Paths
-----

Simulation with a complex differentiation path.

```{r paths}
sim.paths <- splatSimulatePaths(groupCells = c(100, 100, 100),
                                path.from = c(0, 1, 1), de.facLoc = 0.3,
                                de.facScale = 0.6,
                                dropout.present = FALSE,
                                seed = 10)
sim.paths <- plotPCA(sim.paths, return_SCESet = TRUE, draw_plot = FALSE)
```

```{r paths-PCA}
pca.paths <- data.frame(PC1 = reducedDimension(sim.paths)[, 1],
           PC2 = reducedDimension(sim.paths)[, 2],
           group = pData(sim.paths)$Group,
           step = pData(sim.paths)$Step) %>%
    ggplot(aes(x = PC1, y = PC2, colour = step, shape = group)) +
    geom_point(size = 4) +
    scale_colour_gradient(low = mcriPalette("themes")[2],
                          high = mcriPalette("themes")[1],
                          name = "Step") +
    scale_shape_manual(name = "Path", values = c(16, 17, 18)) +
    ggtitle("Simulation with paths") +
    theme_cowplot() +
    theme(plot.title = element_text(size = 20, hjust = 0.4),
          axis.text = element_text(size = 12),
          legend.title = element_text(size = 14),
          legend.text = element_text(size = 12),
          legend.key.size = unit(15, "points"),
    )

pca.paths
```

Complex simulations
-------------------

More the paper we want a panel showing the two kinds of complex simulations:
groups and paths.

```{r complex-panel}
panel <- plot_grid(pca.groups, pca.paths, labels = "AUTO")
panel

save_plot("../figures/figure5_complex.pdf", panel,
          ncol = 2, nrow = 1, base_height = 6)
save_plot("../figures/figure5_complex.png", panel,
          ncol = 2, nrow = 1, base_height = 6)
```

Session info
============

```{r session-info, cache = FALSE}
devtools::session_info()
```
