---
title: "Supplementary"
author: "Luke Zappia"
date: '`r Sys.Date()`'
output: html_document
---

Code version: `r system("git log -1 --format=oneline | cut -d' ' -f1", intern = TRUE)`

```{r knitr, include = FALSE}
DOCNAME = "supplementary"
knitr::opts_chunk$set(autodep        = TRUE,
                      cache          = TRUE,
                      cache.path     = paste0("cache/", DOCNAME, "/"),
                      cache.comments = TRUE,
                      echo           = TRUE,
                      error          = FALSE,
                      fig.align      = "center",
                      fig.path       = paste0("figures/", DOCNAME, "/"),
                      fig.width      = 10,
                      fig.height     = 8,
                      dev            = c('svg', 'png'),
                      message        = FALSE,
                      warning        = FALSE)
```

```{r libraries}
# RNA-seq
library("scater")
library("splatter")

# Plotting
library("cowplot")

# Tidyverse
library("tidyverse")
```

```{r source, cache = FALSE}
source("../R/load_datasets.R")
source("../R/utils.R")
```

Introduction
============

In this document we are going to create some supplementary figures.

We start by loading the Tung data which will be as our real dataset.

```{r real}
root <- "/group/bioi1/shared/public_data/scRNA-seq"

datasets <- read_tsv(file.path(root, "datasets.txt"),
                     col_types = cols(.default = col_character(),
                                      NumCells = col_integer()
                                      )
                     )

real <- loadDataset(datasets[7, ], root)

set.seed(1)
real <- real[rowSums(real) > 0, ]
real <- real[, sample(1:ncol(real), 200)]
```

Mean-dropout
============

The first supplementary figure is an example of fitting the logistic function
to the relationship between mean expression the of percentage of zero counts
in a gene. We fit the relationship to the Tung data...

```{r mean-dropout}
# Get library size normalised counts
counts <- real
lib.sizes <- colSums(counts)
lib.med <- median(lib.sizes)
norm.counts <- t(t(counts) / lib.sizes * lib.med)
norm.counts <- norm.counts[rowSums(norm.counts > 0) > 1, ]

# Calculate the means and zeros percentage
means <- rowMeans(norm.counts)
x <- log(means)
obs.zeros <- rowSums(norm.counts == 0)
y <- obs.zeros / ncol(norm.counts)
df <- data.frame(x, y)

# Fit the logistic function
fit <- nls(y ~ logistic(x, x0 = x0, k = k), data = df,
           start = list(x0 = 0, k = -1))
mid <- summary(fit)$coefficients["x0", "Estimate"]
shape <- summary(fit)$coefficients["k", "Estimate"]
```

...then make the plot.

```{r mean-zeros-plot}
gg <- ggplot(df, aes(x = x, y = y)) +
    geom_point() +
    stat_function(fun = logistic, args = list(x0 = mid, k = shape),
                  colour = "red", size = 1) +
    xlab("Mean expression") +
    ylab("Proportion zeros") +
    ggtitle("Fit of mean-zeros relationship") +
    theme_cowplot() +
    theme(plot.title = element_text(size = 20, hjust = 0.4),
          axis.text = element_text(size = 12),
          legend.title = element_text(size = 14),
          legend.text = element_text(size = 12),
          legend.key.size = unit(25, "points"),
    )

gg
ggsave("../output/mean-zeros.png", width = 8, height = 5)
```

Dispersion estimation
=====================

The second supplementary plot shows the correction we have applied to the
dispersion estimates from `edgeR`. Here we simulate data with a variety of
dispersions, use `edgeR` to estimate them then apply the correction.

```{r disps}
# Dummy design matrix
design <- matrix(1, 100, 1)
disps <- sapply(seq(0.01, 2, length.out = 10), function(disp) {
    sim <- splatSimulate(bcv.common = disp, verbose = FALSE, seed = 1)
    est <- edgeR::estimateDisp(counts(sim), design = design)
    return(c(Disp = disp, Est = est$common.dispersion))
})
```

Then we can plot both the raw and correcte estimates against the true simulated
values.

```{r dips-plot}
gg <- disps %>%
    t() %>%
    data.frame() %>%
    rename(edgeR = Est) %>%
    mutate(Corrected = 0.1 + 0.25 * edgeR) %>%
    gather(key = Estimate, value = Value, -Disp) %>%
    mutate(Estimate = factor(Estimate, levels = c("edgeR", "Corrected"))) %>%
    ggplot(aes(x = Disp, y = Value, colour = Estimate)) +
    geom_abline(intercept = 0, slope = 1, colour = "red", size = 1) +
    geom_point(size = 3) +
    scale_color_brewer(palette = "Dark2") +
    xlab("Simulated dispersion") +
    ylab("Estimated dispersion") +
    ggtitle("Dispersion estimates") +
    theme_minimal() +
    theme(plot.title = element_text(size = 24, face = "bold", hjust = 0.4),
          axis.text = element_text(size = 12),
          legend.title = element_text(size = 14),
          legend.text = element_text(size = 12),
          legend.key.size = unit(15, "points")
    )

gg
ggsave("../output/disp_estimates.png", width = 8, height = 5)
```

Non-linear genes
================

Another supplementary figure shows an example of a simulated gene with
non-linear changes in expression over the course of a differentiation
trajectory. We simulate a path and then filter out some of the non-linear genes.

```{r non-linear}
sim <- splatSimulatePaths(de.facLoc = 0.1, de.facScale = 0.4, seed = 1)
cpm(sim) <- edgeR::cpm(counts(sim))

fData(sim) %>%
    filter(DEFacPath1 == 1, SigmaFacPath1 != 0) %>%
    arrange(-GeneMean) %>% head
```

Then we select one to plot.

```{r non-linear-plot}
gene <- "Gene4655"
gg <- data.frame(CPM = assayData(sim)$BaseCellMeans[gene, ], Step = pData(sim)$Step) %>%
    ggplot(aes(x = Step, y = CPM)) +
    geom_point() +
    geom_smooth(se = FALSE) +
    ggtitle("Example non-linear gene") +
    theme_minimal() +
    theme(plot.title = element_text(size = 24, face = "bold", hjust = 0.4),
          axis.text = element_text(size = 12),
          legend.title = element_text(size = 14),
          legend.text = element_text(size = 12),
          legend.key.size = unit(15, "points")
    )

gg
ggsave("../output/nonlinear_gene.png", width = 8, height = 5)
```

Session info
============

```{r session-info, cache = FALSE}
devtools::session_info()
```
